---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## tabulR

[![Linux/OSX](https://travis-ci.org/itsdalmo/tabulR.svg?branch=master)](https://travis-ci.org/itsdalmo/tabulR)
[![Windows](https://ci.appveyor.com/api/projects/status/github/itsdalmo/tabulR?branch=master&svg=true)](https://ci.appveyor.com/project/itsdalmo/tabulR)
[![Coverage](http://codecov.io/github/itsdalmo/tabulR/coverage.svg?branch=master)](http://codecov.io/github/itsdalmo/tabulR?branch=master)

tabulR includes just one function `qtable()`, which is meant to generate a "quick"
table using [data.table](https://github.com/Rdatatable/data.table), and aims to
balance information and readability in the output.

Note: This is a work in progress.

## Installation

#### Dependencies ahead of CRAN

- [data.table](https://github.com/Rdatatable/data.table) 1.9.7: Lets us use `drop = FALSE`
for only RHS of formula in `dcast`.

```R
# Install development version of data.table
install.packages("data.table", type = "source", repos = "https://Rdatatable.github.io/data.table")
```

#### Install tabulR

Development version:
```R
devtools::install_github("itsdalmo/tabulR")
```

CRAN:
```R
# Not on CRAN yet.
```

## Usage

`qtable` takes one or more variables and returns the following, dep. on type:

- `numeric`: The count and weighted mean.
- `factor`/``character`: Count and weighted proportions.
- `date` (including POSIX): Count and Min/Max dates. 

`qtable` always does the following:

- Returns the count for each variable/group (excluding NA's). Multiple counts are separated with **/**.
- Only proportions and means are weighted, the counts are always netural.
- It completes the data to make implicit missing values *explicit*. 
- When completing the data, counts are set to `0`, proportions to `0` and means to `NA`.
- By default, it casts the data to a wide format when deemed appropriate (usually using the last
group in `groups`).
- It also includes a margin (named *Total*) for each column by default.

```{r, message = FALSE}
require(tabulR)
set.seed(100L)
df <- data.frame(
  group = factor(paste("Group", LETTERS[1:3]), levels = paste("Group", LETTERS[1:4])),
  fct = factor(c("No", "Yes", NA), levels = c("Yes", "No", "Don't know")),
  num = runif(3, 0, 100),
  stringsAsFactors = FALSE
)
```

#### Numeric

```{r}
out <- qtable(df, vars = "num", groups = c("group", "fct"))
knitr::knit_print(out, digits = 1L)
```

#### Factor

```{r}
out <- qtable(df, vars = "fct", groups = "group")
knitr::knit_print(out, digits = 1L)
```

