---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## tabulR

[![Linux/OSX](https://travis-ci.org/itsdalmo/tabulR.svg?branch=master)](https://travis-ci.org/itsdalmo/tabulR)
[![Windows](https://ci.appveyor.com/api/projects/status/github/itsdalmo/tabulR?branch=master&svg=true)](https://ci.appveyor.com/project/itsdalmo/tabulR)
[![Coverage](http://codecov.io/github/itsdalmo/tabulR/coverage.svg?branch=master)](http://codecov.io/github/itsdalmo/tabulR?branch=master)

tabulR serves two purposes, to generate "quick" tables using `qtable()` and plots using `line_chart()` and `bar_chart()`. 

Note: This is a work in progress.

## Installation

Development version:
```R
devtools::install_github("itsdalmo/tabulR")
```

CRAN:
```R
# Not on CRAN yet.
```

## Usage

`qtable` takes one or more variables and returns a table (made with [data.table](https://github.com/Rdatatable/data.table)) depending on type:

- `numeric`: The count and weighted mean.
- `factor`/`character`: Count and weighted proportions.
- `date` (including POSIX): Count and Min/Max dates. 

`qtable` (and charts by extension) always does the following:

- Returns the count for each variable/group (excluding NA's). Multiple counts are separated with **/**.
- Only proportions and means are weighted, the counts are always neutral.
- It completes the data to make implicit missing values *explicit*. 
- When completing the data, missing counts, proportions and means are set to `0`, `0` and `NA` respectively.
- By default, tables will be cast to a wide format when deemed appropriate (usually using the last group in `groups`).
- Includes a margin (called *Total*) for each column by default, if the data is grouped.

`bar_chart` and `line_chart` also:

- Maps and aggregates data to generate a plot (made with [ggplot2](https://github.com/hadley/ggplot2)).
- Minimizes the number of legends when mapping groups and variables to aestetics.
- Adds the appropriate geom's, including `geom_text()`.
- Attempts to adjust the axis labels and range to improve readability.

```{r, message = FALSE}
require(tabulR)
set.seed(100L)
df <- data.frame(
  group = factor(rbinom(100, 4, .2), labels = paste("Group", LETTERS[1:4])),
  fct = factor(rbinom(100, 2, .3), labels = c("Yes", "No", "Don't know")),
  dum = factor(rbinom(100, 1, .5), labels = c("Answer A", "Answer B")),
  num1 = runif(100, 0, 100),
  num2 = runif(100, 0, 100),
  stringsAsFactors = FALSE
)
```

#### Numeric

```{r}
qtable_(df, vars = "num1", groups = c("group", "fct"))
```

#### Factor

```{r}
qtable_(df, vars = "fct", groups = "group")
```

#### Chart

```{r ex1, warning = FALSE}
bar_chart_(df, vars = "num1", groups = "group")
```

## NSE

`qtable` is not exactly a NSE version (non standard evaluation) of `qtable_`, it also requires dplyr to use. It expects unquoted arguments for `vars`, and supports `one_of()`, `starts_with()` etc. from dplyr. The same is true for the charts, such as `bar_chart` and `bar_chart_`.

#### Table

```{r, message = FALSE}
require(dplyr)
df %>% group_by(group) %>% qtable(starts_with("num"))
```

#### Chart

```{r ex2, message = FALSE, warning = FALSE}
df %>% group_by(group) %>% bar_chart(starts_with("num"), margin = FALSE)
```
